---
title: "arquivo"
author: "Lara"
date: "2023-07-04"
output: html_document
---

-----------------------------------------------------------------------------------------------
# ME115
# Classroom R exercises!
## I have made up this data to serve as an example for practicing.
-----------------------------------------------------------------------------------------------

# Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)

library(tidyverse)
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(magrittr)
library(lubridate)
library(readr)
library(readxl)
library(knitr)
```

# Directory (relative path)
```{r }
getwd()
enem.original<-read_xlsx("../ME115/enem.xlsx", sheet = "Sheet1")
str(enem.original)
```

# Data cleaning 
```{r }
# Sexo
table(enem.original$Sexo)
enem.original$Sexo %<>% str_squish() %>% str_to_title() %>% str_replace_all(pattern = "^.*Ale.*$|^Male$|^Masculino$", replacement = "M")
enem.original$Sexo %<>% str_squish() %>% str_to_title() %>% str_replace_all(pattern = "^Female$|^Mulher$", replacement = "F")

# DataNasc
enem.original$DataNasc <- as.character(enem.original$DataNasc)
enem.original$DataNasc <- ymd(enem.original$DataNasc)
```

# Create the variable "Age" from DataNasc. You must consider that students performed the exame on 20-Oct-2015.
```{r }
datadaprova <- ymd("2015-10-20")
enem.original$Age <- interval(enem.original$DataNasc, datadaprova)/dyears(1)

View(enem.original)
str(enem.original)
```
# Create a table for Age with the min, max, quantiles, mean, sd values.
```{r }

```
# Plot an histogram for Age. Interpret it.
```{r }

```
# Create the variable FinalGrade, from the mean of CN,CH,LC,MT. Make a list of the best 10 students based on their grades.
```{r }

```
# Create an objetc "enem_jovens" with only the single ones and with age >= 17 and <=24. How many of them were deleted from the database?
```{r }

```






