---
title: "arquivo"
author: "Lara"
date: "2023-07-04"
output: html_document
---

-----------------------------------------------------------------------------------------------
# ME115
# Classroom R exercises!
## I have made up this data to serve as an example for practicing.
-----------------------------------------------------------------------------------------------

# Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)

library(tidyverse)
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(magrittr)
library(lubridate)
library(readr)
library(readxl)
library(knitr)
```

# Directory (relative path)
```{r }
getwd()
enem.original<-read_xlsx("../ME115/enem.xlsx", sheet = "Sheet1")
str(enem.original)
```

# Data cleaning 
```{r }
# Sexo
table(enem.original$Sexo)
enem.original$Sexo %<>% str_squish() %>% str_to_title() %>% str_replace_all(pattern = "^.*Ale.*$|^Male$|^Masculino$", replacement = "M")
enem.original$Sexo %<>% str_squish() %>% str_to_title() %>% str_replace_all(pattern = "^Female$|^Mulher$", replacement = "F")

# DataNasc
enem.original$DataNasc <- as.character(enem.original$DataNasc)
enem.original$DataNasc <- ymd(enem.original$DataNasc)
```

# Create the variable "Age" from DataNasc. You must consider that students performed the exame on 20-Oct-2015.
```{r }
datadaprova <- ymd("2015-10-20")
enem.original$Age <- interval(enem.original$DataNasc, datadaprova)/dyears(1)
table(is.na(enem.original$Age))
View(enem.original)
str(enem.original)
```
# Create a table for Age with the min, max, quantiles, mean, sd values.
```{r }
descritiva <- function(variavel){
  medidas <- summary(variavel)
  standard.deviation <-sd(variavel)
  return(c(medidas, SD = standard.deviation))
}

descritiva <- descritiva(enem.original$Age)
kable(descritiva, col.names = "valor" , digits = 2, caption = "Teste")
```
# Plot an histogram for Age. Interpret it.
Candidates were 15-54 yo. Distribuition is assymetric. The majority is young (mean age  = 21.51)
```{r }
enem.original %>% ggplot(aes(x = Age)) + geom_histogram(binwidth = 0.8, col = 1) +
  labs(x = "Idade dos candidatos", y = "Frquência", title = "Distribuição das idades") +
  theme_bw()

```
# Create the variable FinalGrade, from the mean of CN,CH,LC,MT. Make a list of the best 10 students based on their grades.
```{r }
# enem.original$FinalGrade <- apply(enem.original[, 4:7], 1, mean)
# or
enem.original <- enem.original %>% mutate(FinalGrade = (NotaCN + NotaCH + NotaLC + NotaMT)/4)

#best students:
enem.original %>% slice_max(FinalGrade, n = 10)
```
# Create an object "enem_jovens" with only the single ones and with age >= 17 and <=24. How many of them were deleted from the database?
```{r }
enem_jovens <- enem.original %>% filter(EstadoCivil == "Solteiro") %>% filter(Age >= 17 & Age <= 24)
```
# Create a table with n, mean, sd of FinalGrade by sex and EscolaEM 
```{r }
datayoung <- enem_jovens %>% select(c("Sexo", "EscolaEM", "FinalGrade")) %>% group_by(EscolaEM, Sexo) %>% summarise(n = n(), Media = mean(FinalGrade), standard.deviation = sd(FinalGrade))

kable(datayoung, digits = 1, caption = "Tabela")
```
# Bar graph for Mean - by Sex and EscolaEM 
```{r }
datayoung %>% ggplot(aes(x = EscolaEM, y= Media, fill = Sexo)) + geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Tipo de escola", y = "Media", fill = "Sexo", title = "Gráfico...", subtitle = "Fonte: ficticia") +
  theme_bw()
```

# Boxplot for FinalGrade by InstrMae
```{r }
enem_jovens %>% ggplot(aes(x = InstrMae, y = FinalGrade)) + geom_boxplot() +
  labs(x = "Escolaridade da Mae", y = "Nota final", title = "Gráfico...", subtitle = "Fonte: ficticia") +
  theme_bw()
```
# Boxplot for FinalGrade by InstrMae and EscolaEM
```{r }
enem_jovens %>% ggplot(aes(x = EscolaEM, y = FinalGrade, fill = InstrMae)) + geom_boxplot() +
  labs(x = "Tipo de Escola", y = "Nota final", fill = "Mae", title = "Gráfico...", subtitle = "Fonte: ficticia") +
  theme_bw()
```
# Scatterplot NotaMT and NotaCN. Color by EscolaEM
```{r }
enem_jovens %>% ggplot(aes(x = NotaMT, y = NotaCN, color = EscolaEM)) + geom_point() +
  labs(x = "Mat", y = "Natureza", color = "Tipo Escola", title = "Gráfico...", subtitle = "Fonte: ficticia") +
  theme_bw()
```







